Automated Metadata Generation for Fish Specimen Image Collections {\footnotesize \thanks{Research supported by NSF OAC \#1940233 and \#1940322.} }

### 2021-04-01 paper/paper_new.tex
Minor number tweak.
Grammar

%abstract can be shortened a good bit.
%can offer an alternative manana..
\begin{abstract}
Metadata is a key data source particularly for researchers seeking use [apply] machine learning to the vast collections of digitized specimens. Unfortunately, the metadata available is often sparse and, at times, erroneous; and it is prohibitively expensive to address these limitations through traditional, manual [*human] means. This paper reports on a research applying machine-driven [*automatic] approaches to analyze digitized fish images and extract a variety of important features. The digitized fish specimen are being analyzed as part of the XXX project [*Biology Guided Neural Networks (BGNN)], where researchers are developing a novel class of artificial neural networks using phylogenies and anatomy ontologies. The paper reports on research combining machine  learning  and  image  informatics  tools and techniques. A sample 7247 images from the INHS (write-out) were examined via automatic processes; and results show we can  accurately determine metadata, such as, fish quantity, fish orientation, image scaling based on ruler identification and measurement, and general image quality metrics. Results also indicate that automatic approaches "can outperform humans" [and/or "expedite"] the creation of metadata that's critical to leveraging digitized specimens for machine learning. The sections that follow cover background context, reviews the research design, reports results Next,  the  results  are presented, followed by a discussion. 
%items in square brackets are just other suggested words, e.g., select one or the other, or .... 
%something about implications above
%something about comparison to human results.

% Over the last several decades advances in computing, imaging, and cyberinfrastructure have had a major impact on scientific research and discovery. One area of considerable activity is the digitization of the biological specimens that have been acquired worldwide by museums and other research institutions, undertakings that have produced large image collections.The metadata that is vital for subsequent machine learning, analysis and scientific discovery based on these specimen image repositories is often unavailable, sparse or incorrect. As a step towards improving metadata in specimen research image collections, our team is developing methods for automatically analyzing fish images to extract a variety of important features. These fish specimens are being studied for a larger project entitled Biology Guided Neural Networks (BGNN), which is developing a novel class of artificial neural networks that can exploit the machine readable and predictive knowledge about biology that is available in specimen images, phylogenies and anatomy ontologies. Using a combination of machine learning and image informatics tools and techniques, we can accurately determine metadata such as fish quantity and location within images, fish orientation and other quantitative fish features, image scaling based on ruler identification and measurement, and general image quality metrics for a substantial number of the images being used in the BGNN project.Our goal is to develop image metadata generation methods that both support the research underway within the BGNN project, and provide a framework for future technology developments that can be deployed by repository curators to improve and bolster the metadata they provide with their specimen images. A longer term goal is to extend the image analysis methods for computing quantitative features in support of specific biological investigations.
\end{abstract}

@@

\section{Introduction}
% Moving teaser since this template doesn't actually ask for one
Over the last several decades advances in computing, imaging, and cyberinfrastructure have supported the growth of digital natural history collections, many of which have specimen images.~\cite{beaman2012mass} Additionally, initiatives, such as National Science Foundation’s Advancing Digitization of Biodiversity Collections (ADBC) program, which ran for nearly a decade, have supported the digitization and curation of tens of thousands of biological specimens.~\cite{page2015digitization} These digital specimen accessible through global, open access repositories; they enable researchers, educators, students, and the general public to examine and compare in ways previously unimaginable. Significant challenges surface, however, when researchers seek to apply computational methods to these data items, due to both image quality and metadata issues.
%{jane}should we try to insert machine learning here, last sentence, or computation methods ok?

Unfortunately, potential scientific advances are hindered by image quality
problems and the lack of accurate and pertinent metadata
associated with the image collections.
Poor quality images, e.g. with low contrast, inadequate lighting,
out-of-focus or cluttered visual arrangements, provide inadequate input
for automated
image analysis and machine learning algorithms and lead to inferior
computational results.
In order to perform quantitative morphometric analysis of the specimens,
the physical scale of the images (pixels/inch) is needed; thus
requiring the ability to identify and measure rulers in the images.
Many specimen collections do include Darwin Core metadata,~\cite{biodiv_info_standards} detailing
specimen taxon, geographic location, and several other descriptive aspects.
Additionally, some digitization efforts record technical metadata, detailing imaging specifications. While these types of metadata are helpful for a
human examining several images at a time, they are insufficient for researchers seeking to apply computational methods to examine thousands of images
to determine if, for instance, a specific fish grows to different lengths
in different habitats.

Since the digital collections may each contain tens of thousands of images, producing metadata for each image via a manual process is prohibitively labor-intensive and infeasible. Methods for automatically computing metadata from images are therefore needed to fully exploit biological image repositories for scientific discovery.
As a step towards improving metadata in specimen research image collections, members of Drexel University's Metadata Research Center are developing methods to automatically analyze fish images and
extract a set of data features that provide important metadata about the
digitized specimen.
The research is being conducted as part of the Biology Guided Neural Networks (BGNN) project, which aims to develop a novel class of artificial neural networks that exploit machine readable and predictive knowledge associated with specimen images, phylogenies and anatomy ontologies.
Using a combination of machine learning and image informatics tools and techniques, we can accurately determine general image quality and metadata, such as fish quantity and location within images, fish orientation, and
image scaling based on ruler identification and measurement. Image scaling
allows us to compute quantitative features about the fish specimens, such
as their length and area.  In order to test and validate our methods, they
were  applied to a set of 7,247
images drawn from the Illinois Natural History Survey (INHS) Collection.
The following section of the paper provides contextual background for this work, followed by the research goals and objectives, and a review of our research methods. Next, the results are presented, followed by a discussion. The conclusion highlights key findings and identifies next steps.

\section{Related Work}
\subsection{Metadata for Natural History Image Collection}
%question/comment - we need to keep in mind we are deal with digitized specimens, not simply digital, in other words, the artifacts themselves are digital version of physical, not like an x-ray, where the image is digital from the start.  
A number of different metadata standards have been applied to support the description and access of digitized scientific images. Darwin Core~\cite{biodiv_info_standards} developed specifically for digital museum projects is one of the most frequently used standards. Another approach is to create a metadata application profile by integrating properties from Dublin Core's DCMI Metadata Terms \cite{dc_terms}, the Ecological Metadata Language (EML) \cite{EML_2019} and other namespaces. There is also the recent Audubon Core supporting the discoverability, dissemination, and use of data relating to biological organisms, including digitized specimens.\cite{audub_core} These descriptive standards include metadata properties for taxon, geographic location, and other important specimen content, and have been developed primarily from the eye of a human curator. In other words, their application anticipates that a curator or data entry staff will manually generate metadata, drawing from acquisition logs or original specimen labels. This task is labor intensive, and the metadata associated with each digital rendering is limited, particularly for researchers seeking to apply machine learning, and leverage both the image and the metadata for scientific research.

>>>>

%abstract can be shortened a good bit.
%can offer an alternative manana..
\begin{abstract}
     Over the last several decades advances in computing, imaging, and cyberinfrastructure have had a major impact on scientific research and discovery. One area of considerable activity is the digitization of the biological specimens that have been acquired worldwide by museums and other research institutions, undertakings that have produced large image collections.
The metadata that is vital for subsequent machine learning, analysis and
scientific discovery based on these specimen image repositories
is often unavailable, sparse or incorrect.
As a step towards improving metadata in specimen research image collections, our team is developing methods for automatically analyzing fish images to extract a variety of important features. These fish specimens are being studied for a larger project entitled Biology Guided Neural Networks (BGNN), which is developing a novel class of artificial neural networks that can exploit the machine readable and predictive knowledge about biology that is available in specimen images, phylogenies and anatomy ontologies. Using a combination of machine learning and image informatics tools and techniques, we can accurately determine metadata such as fish quantity and location within images, fish orientation and other quantitative fish features, image scaling based on ruler identification and measurement, and general image quality metrics for a substantial number of the images being used in the BGNN project.
Our goal is to develop image metadata generation methods that both support
the research underway within the BGNN project, and provide a framework for future technology developments that can be deployed by repository curators to improve and bolster the metadata they provide with their specimen images. A longer term goal is to extend the image analysis methods for computing quantitative features in support of specific biological investigations.
\end{abstract}

@@

\section{Introduction}
% Moving teaser since this template doesn't actually ask for one
Over the last several decades advances in computing, imaging, and cyberinfrastructure have supported the growth of digital natural history collections, many of which contain specimen images.~\cite{beaman2012mass} Additionally, initiatives, such as National Science Foundation’s Advancing Digitization of Biodiversity Collections (ADBC) program, which ran for nearly a decade, have supported the digitization and curation of tens of thousands of biological specimens.~\cite{page2015digitization} These digital specimen images are accessible through global, open access repositories; they enable researchers, educators, students, and the general public to examine and compare collections in ways previously unimaginable. Significant challenges surface, however, when researchers seek to apply computational methods to these repositories, due to both image quality and metadata issues.
%{jane}should we try to insert machine learning here, last sentence, or computation methods ok?

Unfortunately, potential scientific advances are hindered by image quality
problems and the lack of accurate and pertinent metadata
associated with the image collections.
Poor quality images---e.g.\ those with low contrast, inadequate lighting,
out-of-focus or cluttered visual arrangements---provide inadequate input
for automated
image analysis and machine learning algorithms, and lead to inferior
computational results.
In order to perform quantitative morphometric analysis of the specimens,
the physical scale of the images (pixels/cm) is needed; thus
requiring the ability to identify and measure rulers in the images.
Many specimen collections do include Darwin Core metadata,~\cite{biodiv_info_standards} detailing
specimen taxon, geographic location, and several other descriptive aspects.
Additionally, some digitization efforts record technical metadata, detailing imaging specifications. While these types of metadata are helpful for a
human examining several images at a time, they are insufficient for researchers seeking to apply computational methods to examine thousands of images
to determine if, for instance, a specific fish grows to different lengths
in different habitats.

Since the digital collections may each contain tens of thousands of images, producing metadata for each image via a manual process is prohibitively labor-intensive and infeasible. Methods for automatically computing metadata from images are therefore needed to fully exploit biological image repositories for scientific discovery.
As a step towards improving metadata in specimen research image collections, members of Drexel University's Metadata Research Center are developing methods to automatically analyze fish images and
extract a set of data features that provide important metadata about the
digitized specimen.
The research is being conducted as part of the Biology Guided Neural Networks (BGNN) project, which aims to develop a novel class of artificial neural networks that exploit machine readable and predictive knowledge associated with specimen images, phylogenies and anatomy ontologies.
Using a combination of machine learning and image informatics tools and techniques, we can accurately determine general image quality and metadata, such as fish quantity and location within images, fish orientation, and
image scaling based on ruler identification and measurement. Image scaling
allows us to compute quantitative features about the fish specimens, such
as their length and area.  In order to test and validate our methods, they
were  applied to a set of \(7,244\)
images drawn from the Illinois Natural History Survey (INHS) Collection.
The following section of the paper provides contextual background for this work, followed by the research goals and objectives, and a review of our research methods. Next, the results are presented, followed by a discussion. The conclusion highlights key findings and identifies next steps.

\section{Related Work}
\subsection{Metadata for Natural History Image Collection}
%question/comment - we need to keep in mind we are deal with digitized specimens, not simply digital, in other words, the artifacts themselves are digital version of physical, not like an x-ray, where the image is digital from the start.  
A number of different metadata standards have been applied to support the description and access of digitized scientific images. Darwin Core~\cite{biodiv_info_standards} developed specifically for digital museum projects is one of the most frequently used standards. Another approach is to create a metadata application profile by integrating properties from Dublin Core's DCMI Metadata Terms~\cite{dc_terms}, the Ecological Metadata Language (EML)~\cite{EML_2019} and other namespaces. There is also the recent Audubon Core supporting the discoverability, dissemination, and use of data relating to biological organisms, including digitized specimens.~\cite{audub_core} These descriptive standards include metadata properties for taxon, geographic location, and other important specimen content, and have been developed primarily from the eye of a human curator. In other words, their application anticipates that a curator or data entry staff will manually generate metadata, drawing from acquisition logs or original specimen labels. This task is labor intensive, and the metadata associated with each digital rendering is limited, particularly for researchers seeking to apply machine learning, and leverage both the image and the metadata for scientific research.
